---
title: "pstat_final_project"
author: "Tiana Curry"
date: "5/22/2020"
output: html_document
---

**Checking uploading options to find the best option**
```{r}
# uploading data to test read.csv vs. read.table
T1 <- read.csv("/Users/Tiana/Desktop/adult.csv", header = FALSE)
T2 <- read.table("/Users/Tiana/Desktop/adult.txt")

head(T1)
head(T2)

dim(T1)
dim(T2)
```
We should use read.csv funciton since the read.table combines columns


**loading data to create dataframe with column names**
```{r}
path <- "/Users/Tiana/Documents/GitHub/pstat126_finalproject/adult.csv" #path to adult file in directory 

ColumnNames <- c("age","workclass","worked","fnlwgt","education","marital-status","occupation","relationship","race","sex","capital-gain","capital-loss","hours-per-week","native-country","income") # creating column names for adult.csv file 

adult_df <- read.csv(path, header = FALSE, col.names = ColumnNames) # creating data frame with adult.csv files and ColumnNames vector

dim(adult_df)  # finding the number of rows and columns in the df 
head(adult_df, 10) # showing first 5 rows of created data frame
```

** Finding missing values labeled as " ?"**
**Converting " ?" to official missing values NA**
```{r}
#filter(adult_df == " ?")  # finding values in df that contain " ?"
adult_df[adult_df == " ?"] <- NA # replacing values == " ?" with NA 
adult_df # showing df to see if there are NA values 
colSums(is.na(adult_df)) # shows with columns contain NA values and how many are in each column 
```


** creating new datframe without missing values NA**
```{r}
CensusAdult_df <- na.omit(adult_df) # creating new df without rows that contain NA vlaues 
dim(CensusAdult_df)  # finding the number of rows and columns in the dataframe 
colSums(is.na(CensusAdult_df)) # checking which columns have NA, there shouldn't be anyvalues with NA 
```

```{r}
export_path <- "/Users/Tiana/Documents/GitHub/pstat126_finalproject/CensusAdult_df.csv"
write.csv(CensusAdult_df, export_path, row.names = FALSE)
```
**Resources**
UCI Machine Leanring Repository Adult data set - https://archive.ics.uci.edu/ml/datasets/Adult
Dealing with missing values - https://uc-r.github.io/missing_values
Replacing values with NA - https://stackoverflow.com/questions/11036989/replace-all-0-values-to-na
missing data - https://www.statmethods.net/input/missingdata.html
Data cleaning R and the Tidyverse - https://towardsdatascience.com/data-cleaning-with-r-and-the-tidyverse-detecting-missing-values-ea23c519bc62
filtering Rows that Contain a certian string using R - https://intellipaat.com/community/7401/filtering-row-which-contains-a-certain-string-using-dplyr
Markdown cheat sheet - https://www.markdownguide.org/cheat-sheet/
