---
title: "final_project"
author: "Amelia Meyer"
date: "6/3/2020"
output: pdf_document
---

```{r}
# adding the new dataset without the missing values
adult <- read.csv('CensusAdult_df.csv', header = T)
names(adult)
attach(adult)
```

Names of the variables:
age	workclass	worked	fnlwgt	education	marital.status	occupation	relationship	race	sex	capital.gain	capital.loss	hours.per.week	native.country	income

We chose our response value as capital.gain

Optimal model based on forward selection with AIC: capital.gain ~ income + fnlwgt + capital.loss + workclass + age + relationship + hours.per.week

Our predictor values:
x1:income
x2:fnlwgt
x3:capital.loss
x4:workclass
x5:age
x6:relationship
x7:hours.per.week
```{r}
# Finding predictors (refer to days 24 and 26)
# Using forward selection (bottom up method) with AIC
mod.0 <- lm(capital.gain ~ 1, data = adult) 
mod.full <- lm(capital.gain ~., data = adult) # including all predictors in lm()
step(mod.0, scope = list(lower = mod.0, upper = mod.full), direction = "forward") #Uses AIC by default
```

Optimal model using forward selection with BIC:
capital.gain~income + fnlwgt + capital.loss + workclass + age

Our predictors:
x1:income
x2:fnlwgt
x3:capital.loss
x4:workclass
x5:age
```{r}
# Finding predictors (see days 24 and 26)
# Using forward selection (bottom up method) with BIC
## step() function will still print out "AIC", but in fact it is using BIC
n <- length(adult$capital.gain) # sample size
step(mod.0, scope = list(lower = mod.0, upper = mod.full), direction = 'forward',k = log(n), trace = 0)
```


Optimal model using backward selection with AIC:
capital.gain~age + workclass + fnlwgt + relationship + capital.loss + hours.per.week + income

Our predictors:
x1:age
x2:workclass
x3:fnlwgt
x4:relationship
x5:capital.loss
x6:hours.per.week
x7:income

```{r}
# Finding predictors
# Using backward selection (top down method) with AIC
step(mod.full, scope = list(lower = mod.0, upper = mod.full), direction = 'backward')
```

Optimal model using backward selection with BIC:
capital.gain ~ age + workclass + fnlwgt + capital.loss + income

Our predictors: 
x1: age
x2: workclass
x3: fnlwgt
x4: capital.loss
x5: income

```{r}
# Finding predictors
# Using backward selection (top down method) with BIC
step(mod.full, scope = list(lower = mod.0, upper = mod.full), direction = 'backward',k = log(n), trace = 0)
```

Optimal model using stepwise selection:
capital.gain ~ age + workclass + fnlwgt + relationship + capital.loss + hours.per.week + income

Our predictors:
x1: age
x2: workclass
x3: fnlwgt
x4: relationship
x5: capital.loss
x6: hours.per.week
x7: income

```{r}
# Finding predictors
# Using stepwise selection (hybrid model)
step(mod.0, scope = list(lower = mod.0, upper = mod.full))
step(mod.full, scope = list(lower = mod.0, upper = mod.full))
```

Optimal model using regsubsets:
Based on rsq, our model should contain 2 predictors. Based on adjr2, our model should contain 8 predictors. Based on Cp, our model should contain 7 predictors. Based on BIC, our model should contain 6 predictors. We need to do further analysis in order to determine whether the 7th and 8th predictors are helpful. 

```{r}
library(leaps)
mod.reg <- regsubsets(cbind(age,workclass,worked,fnlwgt,education,marital.status,occupation,relationship,race,sex,capital.loss,hours.per.week,native.country,income),capital.gain, data=adult)
summary.reg <- summary(mod.reg)
names(summary.reg)
summary.reg$which
summary.reg$rsq
summary.reg$adjr2
summary.reg$cp
summary.reg$bic

par(mfrow = c(2, 2))
plot(summary.reg$rsq, xlab = "Number of Variables", ylab = "RSq", type = "b")

plot(summary.reg$adjr2, xlab = "Number of Variables", ylab = "Adjusted RSq", type = "b")
best_adj_r2 = which.max(summary.reg$adjr2)
points(best_adj_r2, summary.reg$adjr2[best_adj_r2],
       col = "red",cex = 2, pch = 20)

plot(summary.reg$cp, xlab = "Number of Variables", ylab = "Cp", type = 'b')
best_cp = which.min(summary.reg$cp[-c(length(summary.reg$cp))])
points(best_cp, summary.reg$cp[best_cp],
       col = "red", cex = 2, pch = 20)

plot(summary.reg$bic, xlab = "Number of Variables", ylab = "BIC", type = 'b')
best_bic = which.min(summary.reg$bic)
points(best_bic, summary.reg$bic[best_bic],
       col = "red", cex = 2, pch = 20)

mod.full <- lm(capital.gain ~., data = adult) 
mod.red <- lm(capital.gain ~income + fnlwgt + capital.loss + workclass + age)
anova(mod.red, mod.full)
```

Scatterplots and avplots

```{r}
# looking at the relationships 
pairs(~capital.gain + income + fnlwgt + capital.loss + workclass + age)

library(car)
scatterplotMatrix(~capital.gain + income + fnlwgt + capital.loss + workclass + age)
```

avPlots to test if we should add B7, marital.status
The output should be linear in order for the predictor that we're thinking about adding to contain useful information in predicting Y

```{r}
# checking to see if predictors 'marital.status' and 'relationship' are necessary for our model

# Running test for B7=0 aka marital.status=0
# Added Varaible Plots
full.lm <- lm(capital.gain ~ income + fnlwgt + capital.loss + workclass + age + marital.status)
avPlots(full.lm, id=FALSE)

# summary table 
summary(full.lm)

# Global F-test, $H_0: \beta_1 = \beta_2 = \beta_3 = 0$
red.lm <- lm(capital.gain ~ 1)
anova(red.lm, full.lm)

# Global F-test, $H_0: \beta_1 = \beta_2 = \beta_3 = 0$ ( ? )
anova(full.lm, red.lm)
```

avPlots to test if we should add B8, relationship
The output should be linear in order for the predictor that we're thinking about adding to contain useful information in predicting Y

```{r}
# Running test for B8=0 aka relationship
# First figure out the output of B7=0
# Added Variable Plots
# full2.lm <- lm(capital.gain ~ income + fnlwgt + capital.loss + workclass + age + relationship) or + marital.status + relationship
# we know marital.status would go before relationship because that was in the output of the regsubsets function
# avPlots(full2.lm, id=FALSE)

# summary table 
# summary(full2.lm)

# Global F-test, $H_0: \beta_1 = \beta_2 = \beta_3 = 0$
# red.lm <- lm(ACL ~ 1)
# anova(red.lm, full2.lm)

# Global F-test, $H_0: \beta_1 = \beta_2 = \beta_3 = 0$ ( ? )
# anova(full2.lm, red.lm)
```


# avPlots to test whether or not we should include marital.status and relationship, do a partial F test to test the hypotheses 
# when you're checking the pairs and scatterplotMatrices, LINE
# You don't necessarily have to do B7=0 and B8=0, you can do B7=B8=0
# Transformation for Multiple Linear Regression using powerTransform(), but don't transform categorical variables
# if Rd Power = 0, log transformation
# Once you have found the optimal model, check for outliers and high leverage points 
# remove any outliers or high leverage points but if we do, mention that we did so because it'll change the model
# get the intercept value and the value of each slope parameter so that we can evaluate the relationship with concrete data 
# predict some x value, create a prediction interval
# confidence interval 


